\Large Generative Dynamics of Supreme Court Citations: \\ Analysis with a New Statistical Model

### 2020-05-04 Tex/paper.tex
minor tergm edit
revised abstract

\author{%
  Christian S. Schmid\footnote{Department of Statistics, The Pennsylvania State University, schmid@psu.edu}%
  \and Ted Hsuan Yun Chen \footnote{Department of Political Science, The Pennsylvania State University, thc126@psu.edu}%
  \and Bruce A. Desmarais \footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
 % \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
  }

\maketitle
\begin{abstract}
\noindent The significance and influence of US Supreme Court majority opinions derive in large part from opinions' roles as precedents for future opinions. A growing body of literature seeks to understand what drives the use of opinions as precedents through the study of Supreme Court case citation patterns. We raise two limitations of existing work on Supreme Court citations. First, dyadic citations are typically aggregated to the case level before they are analyzed. Second, citations are treated as if they arise independent of one another. We make the case that studying Supreme Court citations as a dynamic citation network overcomes both of these limitations. We present a methodology for studying citations between US Supreme Court opinions at the dyadic level, as a network. This methodology---the citation temporal exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation. We apply this methodology to a network that includes all Supreme Court cases decided between 1950 and 2015. We find evidence for dependence processes, including reciprocity, transitivity, and popularity. The dependence effects that we identify are as substantively and statistically significant as the effects of the exogenous covariates we include in the model. The summary result from this analysis is that theoretical models of Supreme Court citation formation should consider both the effects of case characteristics and the structure of past citations.\footnote{
%Prepared for presentation at the 2018 Annual Meeting of the American Political Science Association. 
This work was supported in part by NIH, 
R01 AI36664-01. NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219. We would like to thank Rachael Hinkle and Benjamin Kassow for detailed comments on an earlier draft of this paper.}
\end{abstract}

@@

% Network Statistics
The c-TERGM is specified through 
the decision regarding which network statistics $h(\cdot)$ to incorporate. We include the following statistics for the Supreme Court citation network:
$$h_{edges}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{i=1}^Nc_i(t)$$
the number of citations made at time $t$.  $h_{edges}$ is like an intercept, and models the expected value of any given edge.
$$h_{outstar}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
the number of weighted outstars occuring at time $t$. We argue that it should be more likely to cite more recent cases than cases that have been decided further in the past. For the weight 
$$w(a,b):= \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
we define $a$ and $b$ as the elapsed time since case $i$ and $j$ have been introduced to the network.
$$h_{triangle}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot
c_j(t_{-i}) \cdot w(a,b)$$
where $c_j(t_{-i})$ indicates whether case $j$ was cited at the time case $i$ was introduced into the network. Just as for the outstar statistic, we include a weighting factor to favor more recent cases. \\[0.3cm]

>>>>

\author{%
  Christian S. Schmid\footnote{Department of Statistics, The Pennsylvania State University, schmid@psu.edu}%
  \and Ted Hsuan Yun Chen \footnote{Department of Political Science, The Pennsylvania State University, thc126@psu.edu}%
   \and Bruce A. Desmarais \footnote{Department of Political Science, The Pennsylvania State University, bdesmarais@psu.edu}%
  \and David R. Hunter \footnote{Department of Statistics, The Pennsylvania State University, dhunter@stata.psu.edu}%
  }

\maketitle
\begin{abstract}
\noindent The significance and influence of US Supreme Court majority opinions derive in large part from opinions' roles as precedents for future opinions. A growing body of literature seeks to understand what drives the use of opinions as precedents through the study of Supreme Court case citation patterns. We raise two limitations of existing work on Supreme Court citations. First, dyadic citations are typically aggregated to the case level before they are analyzed. Second, citations are treated as if they arise independently. We present a methodology for studying citations between Supreme Court opinions at the dyadic level, as a network, that overcomes these limitations. This methodology---the citation exponential random graph model---enables researchers to account for the effects of case characteristics and complex forms of network dependence in citation formation. We then analyze a network that includes all Supreme Court cases decided between 1950 and 2015. We find evidence for dependence processes, including reciprocity, transitivity, and popularity. The dependence effects are as substantively and statistically significant as the effects of the exogenous covariates we include in the model, indicating that models of Supreme Court citation should incorporate both the effects of case characteristics and the structure of past citations.\footnote{
Prepared for presentation at the 2018 Annual Meeting of the American Political Science Association. 
This work was supported in part by NIH, 
R01 AI36664-01. NSF grants SES- 1558661, SES-1637089, SES-1619644, and CISE-1320219. We would like to thank Rachael Hinkle and Benjamin Kassow for detailed comments on an earlier draft of this paper.}
\end{abstract}

@@

% Network Statistics
The c-ERGM is specified through 
the decision regarding which network statistics $h(\cdot)$ to incorporate. We include the following statistics for the Supreme Court citation network:
$$h_{edges}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{i=1}^Nc_i(t)$$
the number of citations made at time $t$.  $h_{edges}$ is like an intercept, and models the expected value of any given edge.
$$h_{outstar}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
the number of weighted outstars occuring at time $t$. We argue that it should be more likely to cite more recent cases than cases that have been decided further in the past. For the weight 
$$w(a,b):= \sqrt{\dfrac{(t-a)(t-a-b)}{t^2}}$$
we define $a$ and $b$ as the elapsed time since case $i$ and $j$ have been introduced to the network.
$$h_{triangle}:\mathcal{C}(N)\to \mathbb{R}~~~, ~~~c(t) \to \sum_{j<i}^Nc_i(t)\cdot c_j(t) \cdot
c_j(t_{-i}) \cdot w(a,b)$$
where $c_j(t_{-i})$ indicates whether case $j$ was cited at the time case $i$ was introduced into the network. Just as for the outstar statistic, we include a weighting factor to favor more recent cases. \\[0.3cm]
